\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{setspace}
\usepackage{hyperref}

\setstretch{1.1}

\title{\textbf{STA380 Project Proposal}\\
Monte Carlo Study of the Bias--Variance Tradeoff in Regression Models}

\author{
Jizheng Huang \\
\texttt{jizheng.huang@mail.utoronto.ca}
\and
Victor Jiang \\
\texttt{vic.jiang@mail.utoronto.ca}
\and
Tianchen Xu \\
\texttt{tianchen.xu@mail.utoronto.ca}
\and
Kai Rui Zhu \\
\texttt{kairui.zhu@mail.utoronto.ca}
}

\date{University of Toronto, Mississauga}

\begin{document}
\maketitle

% ---------------------------
\section{Project Topic}
Monte Carlo Study of the Bias--Variance Tradeoff in Regression Models.


% ---------------------------
\section{Simulation vs.\ Dataset}
This project will use pure simulation rather than a real-world dataset.

Data will be generated according to the model
\[
y = f(x) + \varepsilon,
\]
where $f(x)$ is a fixed nonlinear regression function and
\[
\varepsilon \sim \mathcal{N}(0, \sigma^2).
\]
The parameter $\sigma$ represents the \emph{standard deviation} of the Gaussian noise and will be referred to as the
\emph{noise level} throughout the project.

% ---------------------------
\section{Project Details}
Monte Carlo simulation will be used to repeatedly generate independent training and test datasets. For each Monte Carlo
replication, we will:
\begin{enumerate}
    \item generate a training dataset and an independent test dataset from the data-generating process,
    \item fit polynomial regression to the training data, and
    \item evaluate predictive performance on the test data.
\end{enumerate}

Model complexity will be controlled by the polynomial degree $d$. For each degree $d$, we will compute:
\begin{itemize}
    \item training mean squared error (MSE),
    \item test mean squared error (MSE),
    \item Monte Carlo estimates of bias$^2$ and variance (based on repeated fits across simulated datasets).
\end{itemize}

Bias and variance will be estimated by comparing the fitted model's predictions to the true regression function $f(x)$
across a grid of input values, and summarizing the systematic deviation (bias) and variability (variance) across Monte
Carlo replications.


% ---------------------------
\section{User Inputs (Shiny Components)}
Users will be able to modify at least the following interactive inputs in the R Shiny app:
\begin{enumerate}
    \item Random seed (for reproducibility)
    \item Sample size $n$
    \item Noise level $\sigma$ (standard deviation of $\varepsilon \sim \mathcal{N}(0, \sigma^2)$)
    \item Polynomial degree $d$
    \item Choice of true function $f(x)$. Users will select from a set of pre-defined functions to vary the complexity of the target:
    \begin{itemize}
        \item \textbf{Smooth:} $f(x) = \sin(x)$ (low frequency)
        \item \textbf{Wiggly:} $f(x) = \sin(x) + 2\cos(5x)$ (high frequency)
        \item \textbf{Localized:} $f(x) = \exp(-x^2)$ (Gaussian bump)
    \end{itemize}
    \item Plot customization options (e.g., colour/theme toggles)
    \item Option to download the simulated dataset and/or summary results as a \texttt{.csv} file
\end{enumerate}



% ---------------------------
\section{Generative AI Usage Statement}
We used generative AI tools to assist in writing the LaTeX code for the proposal. All statistical methods, project design decisions, and final content were reviewed and validated by team members. We did not use any generative AI tools to generate results, simulations, or analyses.

% ---------------------------

\bibliographystyle{plain}
\nocite{*}
\bibliography{references}

\end{document}
